---
title: "Configuring lychee checks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Configuring lychee checks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(crayon.enabled = FALSE, cli.num_colors = 1)
```

```{r setup}
library(rambutan)
```

rambutan's defaults are usually fine for a quick check, but real projects tend
to need at least one exception: a flaky internal host, a longer timeout for a
slow API, or a status code that should count as "fine, actually". This
vignette covers the two ways to configure `check_url()`, `check_urls()`,
`check_paths()`, and `check_package()`, and how they combine.

## Two ways to configure

- **`lychee_options()`** -- pass an options object to any `check_*()` call.
  Good for one-off overrides, or for settings that differ between calls.
- **`lychee.toml`** -- drop a `lychee.toml` file in your working directory
  and every `check_*()` call picks it up automatically. Good for settings
  you want to share between an interactive session and CI, or for a project
  that already has a `lychee.toml` for the
  ['lychee' CLI](https://github.com/lycheeverse/lychee) -- rambutan reads
  the same file, using the same field names.

Every argument to `lychee_options()` shares its name with the matching
`lychee.toml` field, so the two are interchangeable once you know one of
them.

## Precedence

When both a `lychee.toml` file and a `lychee_options()` argument set the
same thing:

- For most fields (`timeout`, `max_retries`, `accept`, `user_agent`, ...),
  whatever you pass to `lychee_options()` wins over the file, and the file
  wins over lychee-lib's own built-in default.
- For `exclude`, `include`, and `header`, the two sources are **combined**
  rather than one replacing the other -- a URL excluded by either the file
  or the function call is excluded.

The example below writes a `lychee.toml` that excludes port `9001`, then
calls `check_urls()` with an `exclude` for port `9002`. Both URLs end up
excluded, which is only possible if the file and the argument were merged
rather than one overriding the other. Nothing here touches the network --
`9001`/`9002` on localhost are never listened on, so the check is
deterministic and offline-safe.

```{r}
project_dir <- withr::local_tempdir()
writeLines('exclude = ["9001"]', file.path(project_dir, "lychee.toml"))

withr::with_dir(project_dir, {
  check_urls(
    c("http://localhost:9001/aaa", "http://localhost:9002/bbb"),
    options = lychee_options(exclude = "9002")
  )
})
```

Both rows report `is_success = FALSE` with `details` explaining they were
excluded -- not that the connection failed -- confirming the pattern from
the file and the pattern from `lychee_options()` were both applied.

## Configuring per call with `lychee_options()`

```{r, eval = FALSE}
check_urls(
  c(
    "https://www.r-project.org",
    "https://api.example.com/flaky-endpoint"
  ),
  options = lychee_options(
    timeout = 30,
    max_retries = 5,
    accept = c("200..=204", "429"),
    user_agent = "rambutan-ci"
  )
)
```

`lychee_options()` validates its arguments as it builds the object, so a
typo like passing a vector where a single value is expected fails
immediately rather than surfacing a confusing error from inside lychee:

```{r, error = TRUE}
lychee_options(timeout = c(10, 20))
```

## Sharing settings with a `lychee.toml` file

If you already run the `lychee` CLI in CI (for example via
`lycheeverse/lychee-action`), rambutan can read the exact same
`lychee.toml` -- no duplication needed:

```{r}
lychee_toml <- '
exclude = ["\\\\.local", "^https://internal\\\\."]
timeout = 20
max_retries = 3
accept = ["200..=204", "429"]

[header]
Accept = "text/html"
'
writeLines(lychee_toml, file.path(project_dir, "lychee.toml"))
```

With this file in the working directory, a plain call now runs with those
settings applied automatically -- no `options` argument needed:

```{r}
withr::with_dir(project_dir, {
  check_urls(c("http://svc.local", "https://internal.example.com"))
})
```

In your own project this looks even simpler, since you likely won't need to
juggle a temporary working directory:

```{r, eval = FALSE}
check_package(".")
```

Options rambutan doesn't (yet) support are simply ignored rather than
causing an error, so a `lychee.toml` written for the full `lychee` CLI
keeps working here even if it sets something rambutan doesn't expose.

## Reference

See `?lychee_options` for the full list of supported fields --
`exclude`/`include`, `timeout`, `max_redirects`, `max_retries`,
`retry_wait_time`, `user_agent`, `method`, `accept`, the
`exclude_all_private`/`exclude_private`/`exclude_link_local`/
`exclude_loopback` IP filters, `require_https`, `include_mail`, and
`header`.
